---
title: "Final Project"
due date: "12/14/2021"
author: "Gus Vu & Jeremy Bednar"
output: html_notebook
---

### Load Libraries & Loading Both Data Sources

```{r}
library(jsonlite)
library(lubridate)
library(gghighlight)
library(spotifyr)
library(tidyverse)
library(knitr)
library(ggplot2)
library(plotly)
library(esquisse)
library(dplyr)
library(purrr)
library(ggridges)
library(reticulate)

# Gus history
streamHistoryGus0 <- fromJSON("GusStreamingHistory0.json", flatten = TRUE)
streamHistoryGus1 <- fromJSON("GusStreamingHistory1.json", flatten = TRUE)

# Jeremy history
streamHistoryJeremy0 <- fromJSON("JeremyStreamingHistory0 copy.json", flatten = TRUE)
streamHistoryJeremy1 <- fromJSON("JeremyStreamingHistory1 copy.json", flatten = TRUE)
streamHistoryJeremy2 <- fromJSON("JeremyStreamingHistory2 copy.json", flatten = TRUE)
streamHistoryJeremy3 <- fromJSON("JeremyStreamingHistory3 copy.json", flatten = TRUE)
streamHistoryJeremy4 <- fromJSON("JeremyStreamingHistory4 copy.json", flatten = TRUE)

# Joining Data Frames Per User + converting "msPlayed" to seconds
streamHistoryJeremy <-
  streamHistoryJeremy0 %>%
  full_join(streamHistoryJeremy1) %>%
  full_join(streamHistoryJeremy2) %>%
  full_join(streamHistoryJeremy3) %>%
  full_join(streamHistoryJeremy4) %>%
  mutate(Listener = "Jeremy") %>%
  mutate(sPlayed = msPlayed/1000)

streamHistoryGus <-
  streamHistoryGus0 %>%
  full_join(streamHistoryGus1) %>%
  mutate(Listener = "Gus") %>%
  mutate(sPlayed = msPlayed/1000)

#API Access (Gus)
Sys.setenv(SPOTIFY_CLIENT_ID = '949891eb08fe47cb8e6e031a841b76ba')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '54b9813cd7bc40eca15514be0b52701b')

access_token <- get_spotify_access_token()
```

### Research Question
Essentially, we want to recreate Spotify Wrapped in our own way. One question we might ask is, "How much time have I spent listening to music this year?" or, "How has my music taste changed over time?". With our data set, we could answer these questions and many more.

### Where/How Data was Found
```{r}
Link1 <- "https://www.spotify.com/us/account/privacy/"
Link2 <- "https://towardsdatascience.com/get-your-spotify-streaming-histo ry-with-python-d5a208bbcbd3"
```
We knew that Spotify collects the data necessary for a project like this because of their Spotify Wrapped every year. After some googling, we found out that it can be requested and used personally.


### Examine the data source
The data is collected by Spotify. Cases should come in the form of listening instances based on log information. Therefore, the data will show listening time intervals as well as many other variables pertaining to the songs that have been listened to. These variables include but are not limited to...
acousticness — how acoustic,
danceability — self-explanatory,
energy — how 'fast, loud an noisy,
instrumentalness — the less vocals, the higher,
liveness — whether there is audience in the recording,
loudness — self-explanatory,
speechiness — the more spoken words, the higher,
valence — whether the track sounds happy or sad, and
tempo — the bpm.
Spotify created this data for marketing but it is available to users upon request.

## Analysis

### Total Listening Length Per User (in seconds)

```{r}
#Total listening time of the time interval for which the data was collected (including "banned artists")
streamHistoryTotalJeremy <- 
  streamHistoryJeremy %>%
  summarise(TotalPlayed = sum(sPlayed))
streamHistoryTotalJeremy

streamHistoryTotalGus <- 
  streamHistoryGus %>%
  summarise(TotalPlayed = sum(sPlayed))
streamHistoryTotalGus

```

### Listening Length Per User By Song

```{r}
#total listening time per song
streamHistoryJeremySong <- 
    streamHistoryJeremy %>%
    group_by(trackName, artistName) %>%
    select(artistName, trackName, sPlayed) %>%
    summarize(sTotal = sum(sPlayed)) %>%
    mutate(Listener = "Jeremy")

streamHistoryGusSong <- 
    streamHistoryGus %>%
    group_by(trackName, artistName) %>%
    select(artistName, trackName, sPlayed) %>%
    summarize(sTotal = sum(sPlayed)) %>%
    mutate(Listener = "Gus")

```

### Listening Length Per User By Artist

```{r}
#total listening time for all artists
streamHistoryJeremyArtist <- 
    streamHistoryJeremy %>%
    group_by(artistName) %>%
    select(artistName, trackName, sPlayed) %>%
    summarize(sTotal = sum(sPlayed)) %>%
    mutate(Listener = "Jeremy")

streamHistoryGusArtist <- 
    streamHistoryGus %>%
    group_by(artistName) %>%
    select(artistName, trackName, sPlayed) %>%
    summarize(sTotal = sum(sPlayed)) %>%
    mutate(Listener = "Gus")

```

### Gus and Jeremy Full Join (streamHistory, Song, and Artist)

```{r}
#combining the original streamHistory tables for Gus & Jeremy
streamHistoryGusAndJeremyOG <-
  streamHistoryGus %>%
  full_join(streamHistoryJeremy) 

#total listening time per song for both users
streamHistoryGusAndJeremySong<-
  streamHistoryGusSong %>%
  full_join(streamHistoryJeremySong) %>%
  filter(trackName != "Unknown Track") %>%
  arrange(desc(sTotal))

#total listening time for all artists either user listened to
streamHistoryGusAndJeremyArtist <-
  streamHistoryGusArtist %>%
  full_join(streamHistoryJeremyArtist) %>%
  arrange(desc(sTotal))

streamHistoryGusAndJeremySong %>%
  head(10)


```

### Top 10 Songs

```{r}

#Top 10 songs Gus
GusTopTenSong <-
  streamHistoryGusSong %>%
  filter(trackName != "Unknown Track") %>%
  arrange(desc(sTotal)) %>%
  head(10)
ggplot(GusTopTenSong) +
  aes(x = reorder(trackName,+sTotal), fill = artistName, weight = sTotal) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(title = "Gus Top 10 Songs") +
  coord_flip() +
  theme_bw() +
  xlab("Track Name") +
  ylab("Total Listened (Seconds)")

#Top 10 Songs Jeremy
JeremyTopTenSong <-
  streamHistoryJeremySong %>%
  filter(trackName != "Unknown Track") %>%
  arrange(desc(sTotal)) %>%
  head(10)
ggplot(JeremyTopTenSong) +
  aes(x = reorder(trackName,+sTotal), fill = artistName, weight = sTotal) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(title = "Jeremy Top 10 Songs") +
  coord_flip() +
  theme_bw() +
  xlab("Track Name") +
  ylab("Total Listened (Seconds)")
```

### Top 10 Artists
```{r}
#graphs and visuals for significant information
GusTopTenArtist <-
  streamHistoryGusArtist %>%
  filter(artistName != "Unknown Artist") %>%
  arrange(desc(sTotal)) %>%
  head(10)
ggplot(GusTopTenArtist) +
  aes(x = reorder(artistName,+sTotal), fill = artistName, weight = sTotal) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(title = "Top 10 Artists For Gus") +
  coord_flip() +
  theme_bw() +
  xlab("Artist") +
  ylab("Total Listened (Seconds)")

JeremyTopTenArtist <-
  streamHistoryJeremyArtist %>%
  filter(artistName != "Unknown Artist") %>%
  arrange(desc(sTotal)) %>%
  head(10)
ggplot(JeremyTopTenArtist) +
  aes(x = reorder(artistName,+sTotal), fill = artistName, weight = sTotal) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(title = "Top 10 Artists For Jeremy") +
  coord_flip() +
  theme_bw() +
  xlab("Artist") +
  ylab("Total Listened (Seconds)")

```

### Initial CoPlaylist

```{r}
#creating a playlist of songs that both users listen to
CoPlaylist <-
  streamHistoryJeremySong %>%
  filter(sTotal <= 600000) %>%
  inner_join(streamHistoryGusSong, by = c("trackName" = "trackName","artistName" = "artistName")) 
CoPlaylist
```

### Top 10 Songs On CoPlaylist

```{r}
#songs each of us listen to a lot that the other has heard before
GusCoPlay <-
  CoPlaylist %>%
  select(trackName,artistName,sTotal.y,Listener.y) %>%
  arrange(desc(sTotal.y)) %>%
  head(15)
ggplot(GusCoPlay) +
  aes(x = reorder(trackName,+sTotal.y), fill = artistName, weight = sTotal.y) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(title = "Gus' Top 15 Songs On CoPlayList") +
  ylab("Total Time (Seconds)") +
  xlab("Song Name") +
  coord_flip() +
  theme_bw()


JeremyCoPlay <-
  CoPlaylist %>%
  select(trackName,artistName,sTotal.x,Listener.x) %>%
  arrange(desc(sTotal.x)) %>%
  head(15)
ggplot(JeremyCoPlay) +
  aes(x = reorder(trackName,+sTotal.x), fill = artistName, weight = sTotal.x) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(title = "Jeremy's Top 15 Songs On CoPlayList") +
  ylab("Total Time (Seconds)") +
  xlab("Song Name") +
  coord_flip() +
  theme_bw()

```


#Audio Features
```{r}
#Find the average score of audio features for top 5 artists
GusfavArtist1 <- get_artist_audio_features(artist= "Ms. Lauryn Hill")
GusfavArtist2 <- get_artist_audio_features(artist= "Erykah Badu")
GusfavArtist3 <- get_artist_audio_features(artist= "2Pac")
GusfavArtist4 <- get_artist_audio_features(artist= "Lil Wayne")
GusfavArtist5 <- get_artist_audio_features(artist= "Jay-Z")
GusTop5Audio <-
  GusfavArtist1 %>%
  full_join(GusfavArtist2)%>%
  full_join(GusfavArtist3)%>%
  full_join(GusfavArtist4)%>%
  full_join(GusfavArtist5)%>%
  summarise(danceability = sum(danceability)/5,energy = sum(energy)/5,speechiness = sum(speechiness)/5,acousticness = sum(acousticness)/5, instrumentalness = sum(instrumentalness)/5) 


#Find the average score of audio features for top 5 artists
JeremyfavArtist1 <- get_artist_audio_features(artist= "Lil Uzi Vert")
JeremyfavArtist2 <- get_artist_audio_features(artist= "Tyler, The Creator")
JeremyfavArtist3 <- get_artist_audio_features(artist= "Future")
JeremyfavArtist4 <- get_artist_audio_features(artist= "Juice WRLD")
JeremyfavArtist5 <- get_artist_audio_features(artist= "Playboi Carti")
JeremyTop5Audio <-
  JeremyfavArtist1 %>%
  full_join(JeremyfavArtist2)%>%
  full_join(JeremyfavArtist3)%>%
  full_join(JeremyfavArtist4)%>%
  full_join(JeremyfavArtist5)%>%
  summarise(danceability = sum(danceability)/5,energy = sum(energy)/5,speechiness = sum(speechiness)/5,acousticness = sum(acousticness)/5, instrumentalness = sum(instrumentalness)/5)

```



```{r}
#Make new data frame so we can graph easier
Audio_Features <- c("danceability", "energy", "speechiness", "acousticness", "instrumentalness")
Score <- c(402.590400, 408.287008, 158.053660, 91.696080, 6.327396)
Gus_Top_5Audio <- data.frame(Audio_Features, Score) 

ggplot(Gus_Top_5Audio, aes(x =Audio_Features, y = Score, fill = Audio_Features)) +
  geom_col() +
  labs(title = "Average Audio Features of Gus' Top 5 Artists")+
  theme_bw() +
  ylim(0,450)


#Make new data frame so we can graph easier
JAudio_Features <- c("danceability", "energy", "speechiness", "acousticness", "instrumentalness")
JScore <- c(208.6264, 178.6905, 65.90458, 51.64297, 6.977631)
Jeremy_Top_5Audio <- data.frame(JAudio_Features, JScore)

ggplot(Jeremy_Top_5Audio, aes(x =JAudio_Features, y = JScore, fill = Audio_Features)) +
  geom_col() +
  labs(title = "Average Audio Features of Jeremy's Top 5 Artists")+
  xlab("Audio_Features") +
  ylab("Score") +
  theme_bw() +
  ylim(0,450)

```



